\documentclass[12pt]{article}
\textwidth=7in
\textheight=9.5in
\topmargin=-1in
\headheight=0in
\headsep=.5in
\hoffset  -.85in

\pagestyle{empty}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{url}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\begin{document}

\begin{center}
{\bf Introduction to Data Science \\ Homework 3: Due Wednesday September 12 at 2:00pm}
\end{center}

\setlength{\unitlength}{1in}

\begin{picture}(6,.1)
\put(0,0) {\line(1,0){6.5}}
\end{picture}

\renewcommand{\arraystretch}{2}

\vskip.25in

\noindent{\bf  {\Large Exercises:} }

\vskip.25in
  \begin{enumerate}
    \item   Go through sections 1 and 2 of the R Programming course in the {\tt swirl} package, then answer the following questions:
    \begin{enumerate}
      \item What does the function {\tt c} do? 
      \item In which package is the {\tt c} function contained? 
      \item Create a vector
        \begin{verbatim}
          v <- c(1,2,3,4)
        \end{verbatim}
        What command would you type to add 3 to each component of the vector? Type your command and write down the result. 
        \item What does the function {\tt getwd} do? 
        \item What does the function {\tt ls} do? 
        \item What is a working directory? 
        \item In RStudio, how can you list all of the files in your working directory? 
        \item Describe what each of the listed functions do, be detailed
        \begin{itemize}
          \item {\tt  list.files}
          \item {\tt  dir.create}
          \item {\tt  file.create}
          \item {\tt  file.info}
          \item {\tt  file.rename}
          \item {\tt  file.copy}
        \end{itemize} 
    \end{enumerate}
    \item Read through section 4 of the article \emph{Excuse me, do you have a moment to talk
about version control?}, then answer the following questions. 
\begin{enumerate}
  \item What is Git?
  \item What is GitHub? 
  \item Why should someone in data science care about these? 
\end{enumerate} 
  \end{enumerate}
\end{document}
